Reproducible Research: Peer Assessment 1
========================================
Author: LI XIANG   Date: Jan 17, 2015

# Loading and preprocessing data

First 5 rows of data look like below:
```{r}
rdata <- read.csv("activity.csv",colClasses = c("numeric", "character", "numeric"))
head(rdata)
```

Convert date field to Data class; interval field to Factor class
```{r}
rdata$date <- as.Date(rdata$date, format = "%Y-%m-%d")
```

The structure of the processed data looks like below:
```{r}
str(rdata)
```

# Mean total no. of steps taken per day

Calculate total steps per day using "aggregate" functions (ignore the missing values here)
```{r}
steps_per_day <- aggregate(steps ~ date, rdata, sum)
```

Plot it:
```{r}
hist(steps_per_day$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
rmean <- mean(steps_per_day$steps)
rmedian <- median(steps_per_day$steps)
```

The mean is `r rmean` and the median is `r rmedian`.

# Avg daily actvitity pattern

Find out avg steps per interval
```{r}
steps_by_interval <- aggregate(steps ~ interval, rdata, mean)
```

Plot the avg num of steps by interval
```{r}
plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps")
```

Find interval w/ most avg steps
```{r}
int_max <- steps_by_interval[which.max(steps_by_interval$steps),1]
```
Interval with most average steps is `r int_max`.

# Imputing missing values

Calculate and report the total number of missing values in the dataset, then *Missing values* were imputed by inserting the average for each interval.
```{r}
incomplete <- sum(!complete.cases(rdata))
imputed_data <- transform(rdata, steps = ifelse(is.na(rdata$steps), steps_by_interval$steps[match(rdata$interval, steps_by_interval$interval)], rdata$steps))
```

Recount total steps by day
```{r}
steps_by_day_new <- aggregate(steps ~ date, imputed_data, sum)
hist(steps_by_day_new$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")

hist(steps_per_day$steps, main = paste("Total Steps Each Day"), col="red", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "red"), lwd=10)
```

Calculate new statistics
```{r}
rmean.i <- mean(steps_by_day_new$steps)
rmedian.i <- median(steps_by_day_new$steps)
mean_diff <- rmean.i - rmean
med_diff <- rmedian.i - rmedian
total_diff <- sum(steps_by_day_new$steps) - sum(steps_per_day$steps)
```

1. The imputed data mean is `r rmean.i`.
2. The imputed data median is `r rmedian.i`.
3. The difference between the non-imputed mean and imputed mean is `r mean_diff`.
4. The difference between the non-imputed mean and imputed mean is `r med_diff`.
5. The difference between total number of steps between imputed and non-imputed data is `r total_diff`.

# Difference btw weekdays and weekends

Created a plot to compare and contrast number of steps between the week and weekend.
```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekdays), "Weekday", "Weekend"))

steps_by_interval_new <- aggregate(steps ~ interval + dow, imputed_data, mean)

library(lattice)

xyplot(steps_by_interval_new$steps ~ steps_by_interval_new$interval|steps_by_interval_new$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")
```